---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Michael Wenk"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)
# to do
help(woolyrnq)
help(gas)

```

Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)

```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

```{r, echo=TRUE, cache=TRUE}
# to do
frequency(gold)
frequency(woolyrnq)
frequency(gas)

```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do
which.max(gold)

```

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(mytimeseries, facets = TRUE)

```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries)
```

# Exercise 3

Import some monthly Australian retail data from the foler data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349873A"], frequency = 12, start = c(2000, 5))
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts)
ggAcf(myts)

```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer:
Saisonal: Während die Werte innerhalb des Jahres nur gering schwanken, ist in den Wintermonaten ein erheblicher Anstieg zu verzeichnen.

Zyklik: Die vorhandene Zyklik kann im Seasonal Plot erkannt werden. Ab November ist ein Ansteig erkennbar.

Trend: Über den gesamten Betrachtungszeitraum ist im Mittelwert ein Anstieg zu erkennen.

# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
autoplot(goog) + xlab("trading days") + ylab("stock prices") + ggtitle("Daily closing stock prices")
```


```{r, echo=TRUE, message=FALSE, warning=FALSE}
autoplot(writing)
autoplot(fancy)
autoplot(a10)
autoplot(h02)
```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
ggseasonplot(writing)
ggsubseriesplot(writing)

```

Your answer: Während die Daten für den Rest des Jahres recht konstant sind, findet von Juli bis September ein Einbruch statt.
Der Tiefpunkt wird im August erreicht.
Es gibt keine ungewöhnlichen Jahre.



```{r, echo=TRUE}
# fancy
# to do
ggseasonplot(fancy)
ggsubseriesplot(fancy)

```

Your answer:#
Von Anfang bis Ende des Jahres ist ein Anstieg zu erkennen.
Von November bis Dezember ist der Ansteig am höchsten.
Der Hochpunkt wird im Dezember erreicht.

1990 ist ein relativ schwaches Jahr.


```{r, echo=TRUE}
# a10
# to do
ggseasonplot(a10)
ggsubseriesplot(a10)
```

Your answer: Der Hochpunkt ist im Januar erreicht. Ab da fällt es stark bis April, um sich bis Ende des Jahres wieder zu erholen.

2008 verhält sich ungewöhnlich. Da ist der März tiefer als der Februar.


```{r, echo=TRUE}
# h02
# to do
ggseasonplot(h02)
ggsubseriesplot(h02)
```

Your answer:Der Hochpunkt ist im Januar erreicht. Ab da fällt es stark bis Februar, um sich bis Ende des Jahres wieder zu erholen.

Eine pinke Kurve(2006?) verhält sich ungewöhnlich. Da ist der April tiefer als der Februar.


# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```

Your answer:
+ Seasonality evident in all plots: Hochpunkt im juli, fällt bis Februar, ab da steigt es
+ Cyclicity wird in den ersten beiden Grafiken sichtbar
+ Keine trends


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq)
```

Your answer:
+ Seasonality evident in all plots: Von Quartal 1 steigt es bis Quartal 2. Nach Quartal 3 fällt es.
+ Cyclicity wird in den ersten beiden Grafiken sichtbar
+ trend nach oben erkennbar in grafik 1


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(sunspotarea)
#ggseasonplot(sunspotarea)
ggsubseriesplot(sunspotarea)
gglagplot(sunspotarea)
ggAcf(sunspotarea)
```

Your answer: Daten sind nicht saisonal
kein trend
Zyklic gut erkennbar.


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gasoline)
ggseasonplot(gasoline)
#ggsubseriesplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline)
```

Your answer:
trend nach oben erkennbar in grafik 1
saisonalität: im winter höher als im sommer
zyklik eher unkonstant


# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
set.seed(1711)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
```

```{r}
ggAcf(y)
```


```{r, echo=TRUE, cache=TRUE}
# todo
mypigs <- window(pigs, start=1990)
autoplot(mypigs)
autoplot(y)
ggAcf(mypigs)
ggAcf(y)
```

Your answer:

Ein Muster im Zeitplot ist kaum erkennbar.
Das ACF zeigt eine signifikante Autokorrelation an den Stellen 1,2 und 3
r 12 zeigt eine leichte Saisonalität.
Das Ergebnis ist kein white noise.




# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r}
# todo
ddj <- diff(dj)
autoplot(ddj)
autoplot(y)
ggAcf(ddj)
ggAcf(y)
```

Your answer: Im Zeitplot kann kaum ein Muster erkannt werden. Die Volatilität ist relativ hoch.
Das ACF zeigt keine signifikante Autokorrelation.
An der Stelle r6 wird die untere Grenze knapp unterschritten. Dies ist aber nicht signifikant.

Das Ergebnis scheint ein white noise zu sein..

